---
title: "Week 6 Exercises"
output: html_notebook
---

# Exercise 1

The data frames df_w and df_f represent repeated measures data from 60 participants. Variables F1-F3 and W1-W3 are "sub-variables" that will be used to make two composite variables F_total and W_total, respectively. 

```{r include=F}
# simulating data
set.seed(1)
df_f <- data.frame(id = rep(1:60, each = 2) %>% as.character,
                 session = rep(2:1, 60),
                 group = rep(1:2, each = 60),
                 F1 = sample.int(4, 120, replace = T) - 1,
                 F2 = sample.int(4, 120, replace = T) - 1,
                 F3 = (sample.int(4, 120, replace = T) - 1)) %>%
  mutate(F3 = F3*(session-1), F2 = F2*(group-1)*(2-session)) %>%
  mutate(group = factor(group))

df_w <- data.frame(id = rep(1:60, each = 2) %>% as.character,
                 session = rep(1:2, 60) %>% factor,
                 group = rep(1:2, each = 60) %>% factor,
                 W1 = sample.int(4, 120, replace = T) - 1,
                 W2 = sample.int(4, 120, replace = T) - 1,
                 W3 = sample.int(4, 120, replace = T) - 1)
```

#### 1.1

Merge the two data frames together.

```{r}
df_w # look at df_w data frame
df_f # look at df_f data frame
df <- merge(df_w, df_f, all = TRUE) #merge two data frame together
df # look at the merged data frame
```

#### 1.2

Using the merged data frame, create the **composite variables F_total and W_total, which are the sums of F1-F3 and W1-W3, respectively** (i.e. their values can range from 0 to 9).

```{r}
df %>%
  rowwise() %>%
  mutate(F_total = sum(c(F1, F2, F3))) -> df # create a new variable 'F_total' is the sum of F1, F2 and F3
df %>%
  rowwise() %>%
  mutate(W_total = sum(c(W1, W2, W3))) -> df # create a new variable 'W_total' is the sum of W1, W2 and W3
df # look at the data frame
```


# Exercise 2

#### 2.1

Visualise the distributions of F_total and W_total for the two groups and measurement sessions (for example as boxplots). 

```{r}
df %>%
  ggplot(aes(session, F_total)) + 
  geom_boxplot() +
  facet_wrap(~group)
df %>%
  ggplot(aes(session, W_total)) + 
  geom_boxplot() +
  facet_wrap(~group)
```

#### 2.2

Fit a linear regression model with F_total as the DV, and session and group as predictors.  

```{r}
Model <- lm(F_total ~ session + group, data = df)
summary(Model)
```

#### 2.3

Look at the means of F_total by group and session. How are they linked to the linear regression model coefficients?

```{r}
# look at the mean of 'F_total' by group and session
df_mean <- df %>%
  group_by(group, session) %>%
  summarise(mean_F_total = mean(F_total)) %>%
  ungroup
df_mean
# how the mean link to lm coefficients
  #As the result in linear regression model above, the coefficient of session is not statistically significant, but the coefficient of group is significant. After calculating the mean of F_total by group and session, it has been found that in group1, the mean increase from session1 to session2, while in group2, the mean decrease from session1 to session2. 
```


# Exercise 3

Visualise the [anscombe dataset](https://rdrr.io/r/datasets/anscombe.html) using ggplot2.

```{r}
# Read data
library(datasets)
anscombe
str(anscombe)
summary(anscombe)
# Calculate mean of x1, x2, x3 and x4 to 'x'; calculate mean of y1, y2, y3 and y4 to 'y'
library(tidyverse)
anscombe %>%
 rowwise() %>%
 mutate(x = mean(c(x1, x2, x3, x4))) -> anscombe
anscombe %>%
 rowwise() %>%
 mutate(y = mean(c(y1, y2, y3, y4))) -> anscombe
anscombe
# Visualization (Lm)
anscombe_lm <- lm(y ~ x, data = anscombe)
summary(anscombe_lm)
library(ggplot2)
ggplot(anscombe, aes(x = x, y = y)) +
  geom_point() +
  stat_smooth(method = "lm", col = "red")
```


# Exercise 4

## I will do the alternative final assignment

Prepare a **separate R Notebook/Markdown document**, which will be the first draft of your final assignment with your own data. In the draft, include the following:  

1. Outline the study design, your research question, DV(s), IV(s)  
2. Data wrangling: start from reading in the raw data and show all steps  
3. Plot your main result(s)    

Even if you had already completed some of these steps before, please include all of them in your document. **NOTE: Return either a readable HTML document (.html or .nb.html), or an .Rmd file along with your data, to make it possible for us to review your work!** Make the document as professional-looking as possible (you can, of course, include your comments/questions in the draft). You will get feedback on the draft, based on which you can then make the final version. The final document should be a comprehensive report of your data wrangling process and results.  