---
title: "Week 2 Exercises"
author: "Liyuan E"
date: "September 20th, 2021"
output: 
  html_document:
      toc: true
---

# Exercise 1

#### 1.1

Create a new code chunk where you load the tidyverse package. In the chunk settings, suppress any output messages.  

```{r}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library(tidyverse)
```


#### 1.2

The tibble *df* has 60 observations (rows) of variables (columns) group, gender, score1 and score2 (continuous scores from two tests). Each row represents one participant.  
```{r include = F}
# this chunk generates the data, and you can ignore it
set.seed(1)
df  <- tibble(group = sample(c(rep(1:2, each=30))),
              gender = sample(c(rep(c('M', 'F'), 28), rep(999,4))),
              score1 = rnorm(60, 18, 2),
              score2 = as.character(runif(60, 0.7, 1.1)*score1))
  
```

```{r}
df
```


There is something to fix in three of the variables. Explore the data and describe what needs to be corrected.  

**Hint**: You can use e.g. str(), distinct(), and summary() to explore the data.  
```{r}
# This chunk is to see the summary of the data
df %>% summary()
# The following chunks is to see each variable of the data
df %>%
  distinct(gender)
df %>%
  distinct(score1)
df %>%
  distinct(score2)
# This chunk is to see the structure of the data
df %>% str()
# After exploring the data, the following things need to be corrected: 1. the data do not have id variable which should be added; 2. the missing values in the gender variable coded by 999 which should be re-coded to NA; 3. the group and gender variables need to change to factors; 4. the score2 need to be converted to numerical variable which was done in later exercise.
```


# Exercise 2

#### 2.1

Make the corrections you described above.
```{r}
# recode gender variable 999 to NA (missing)
df <- df %>%
  mutate(gender = na_if(gender, 999))
# insert a new variable "id"
df <- df %>%
  rownames_to_column('sbj_id')
# convert group and gender variables to factors
df <- df %>%
  mutate(group = factor(group), gender = factor(gender), score1 = (score1), score2 = (score2))
```

#### 2.2

Count observations by group and gender. Arrange by the number of observations (ascending).
```{r}
# count observations by group and gender
df %>%
  count(group)
df %>%
  count(gender)
df %>%
  count(group, gender) %>%
# Arrange by the number of observations (ascending)
  arrange(group, gender, (n))
```


# Exercise 3

#### 3.1

Create a new variable, **score_diff**, that contains the difference between score1 and score2.

```{r}
# mutate score2 to numerical variable score2_num
df <- df %>%
  mutate(score2_num = as.numeric(score2))
# create score_diff to contain difference between score1 and score2_num
df$score_diff <- df$score1 - df$score2_num
df %>%
# look score_diff
  distinct(score_diff)
```

#### 3.2

Compute the means of score1, score2, and score_diff.

**Hint**: Like mutate(), summarise() can take multiple variables in one go.

```{r}
# Compute means
df %>%
  summarise(score1_mean = mean(score1),
            score2_mean = mean(score2_num),
            score_diff_mean = mean(score_diff),
            n = n()) %>%
# make mean to 2 decimal places
  mutate(across(is_double, round, 2))
```

#### 3.3

Compute the means of score1, score2, and score_diff **by gender**.

```{r}
grouped_df <- df %>%
  group_by(gender)

grouped_df %>%
  summarise(score1_mean = mean(score1),
            score2_mean = mean(score2_num),
            score_diff_mean = mean(score_diff))
```


# Exercise 4

#### 4.1

Using ggplot2, create a scatter plot with score1 on the x-axis and score2 on the y-axis. 
```{r}
df %>%
  ggplot(aes(x = score1, y = score2)) + 
  geom_point()
```


#### 4.2

Continuing with the previous plot, colour the points based on gender. 

Set the output figure width to 10 and height to 6.


```{r, fig.width= 10, fig.height= 6}
df %>%
  ggplot(aes(x = score1, y = score2)) + 
  geom_point()
```


# Exercise 5

#### 5.1

Add the author (your name) and date into the metadata section. Create a table of contents.  

#### 5.2

Knit your document to HTML by changing `html_notebook` to `html_document` in the metadata, and pressing Knit.   
