---
title: "Week 3 Exercises"
output: html_notebook
---

```{r setup}
# load library
library(data.table)
library(tidyverse)

```

# Exercise 1

#### 1.1

Read the *region_scores.csv* data into R. It should have 240 observations of 6 variables.  

```{r}
# read "region_scores.csv"data
list.files('C:/Users/eliyuan/Documents/GitHub')
df <- readr::read_csv('data/region_scores.csv')
df
```


#### 1.2

Cut the continuous variable **age** into a categorical variable **age_group** that has groups of width 10. Use ggplot2's cutting function.  

Count observations by age group.

```{r}
glimpse(df) 
summary(df)
# cut age to age_group with width 10
df %>%
  mutate(age_group = cut_width(age, 10, boundary = 0)) %>%
count(age_group)
# save age_group
df <- df %>%
  mutate(age_group = cut_width(age, 10, boundary = 0, labels = c('11-20','21-30','31-40','41-50','51-60','61=70'))) 
```


# Exercise 2

The chunk below is supposed to produce a plot but it has some errors.  

The figure should be a scatter plot of **age** on the x-axis and **score** on the y-axis, with points coloured by **education level** (3 levels). It should also have three linear regression lines, one for each of the education levels.  

Fix the code to produce the right figure.  

What happens if you use `geom_jitter()` instead of `geom_point()`?  

**Hint**: Examine the code bit by bit: start by plotting just the scatter plot without `geom_smooth()`, and add the regression lines last.  

```{r}
# variable class transformations
df <- df %>%
  mutate(id = as.character(id),
         region = factor(region),
         education = factor(education, ordered = TRUE),
         gender = factor(gender))
# producing scatter plot that have three regression lines
df %>% 
  ggplot(aes(age, score, color = education)) + 
  geom_point() +
  geom_smooth(method = "lm")

```


# Exercise 3

#### 3.1

Calculate the mean, standard deviation (sd), and number of observations (n) of **score by education and age group**. Also calculate the standard error of the mean (by using sd and n). Save these into a new data frame (or tibble) named edu_age_stats.  

```{r}
# Calculate mean, sd and number of observation (n)
edu_age_stats <- df %>%
  group_by(education, age_group, .drop = FALSE) %>% 
  summarise(mean_score = mean(score),
            sd_score = sd(score),
            n = n()) %>%
  ungroup()

edu_age_stats
 # calculate standard error
edu_age_stats <- edu_age_stats %>%
  mutate(se = sd_score/sqrt(n))
```

#### 3.2

Using edu_age_stats, plot a bar plot that has **age group** on the x-axis, **mean score** on the y-axis, and **education levels** in subplots (facets).  

Use `geom_errorbar()` to add error bars that represent standard errors of the mean.  

```{r}
edu_age_stats %>%
  ggplot(aes(age_group, mean_score)) + 
  geom_col() +  
  facet_wrap(~education) +
  geom_errorbar(aes(ymin = mean_score - se, ymax = mean_score + se))
```


# Exercise 4

#### 4.1

Create a figure that has boxplots of **score** (x-axis) by **region** (y-axis).  

```{r}
# encode region
df <- readr::read_csv('data/region_scores.csv', locale = locale(encoding = "Latin1"))
# check the result
df %>%
  distinct(region)
# create a boxplot
df %>%
  ggplot(aes(x = score, y = region)) + 
  geom_boxplot()
```

#### 4.2

Group the data by region and add mean score by region to a new column **mean_score**. Do this with **`mutate()`** (not `summarise()`). 

Reorder the levels of region based on mean_score.  

**Hint**: Remember to ungroup after creating the mean_score variable.  

```{r}
df <- df %>%
  group_by(region) %>%
  mutate(mean_score = mean(score)) %>% 
  ungroup()
# if use summarise, there will only be two variables: region and mean_score
df
```


# Exercise 5

Using the data you modified in exercise 4.2, plot **mean scores** (x-axis) by **region** (y-axis) as points. The regions should be ordered by mean score.  

Use `stat_summary()` to add error bars that represent standard errors of the mean.  

**Hint**: Be careful which variable - mean_score or score - you're plotting in each of the geoms.  

```{r}
scoreplot <- df %>%
  ggplot(aes(fct_reorder(region, score, mean), score)) + 
  stat_summary(geom = "point", fun.data = "mean_se") +
  stat_summary(geom = "linerange", fun.data = "mean_se") +
  coord_flip() +
  labs(title = "Mean score by region", subtitle = "Lines show standard error of the mean", x = "Region", y = "Mean_score") 

scoreplot
```


